---
title: "Correlation and Simple Linear Regression"
output: html_notebook
---

A. Covariance and correlation
We can compute the covariance and correlation in R using the cov() and cor()
functions.
Ex. A pediatrician wants to study the relationship between a child’s height and
their head circumference (both measured in inches). She selects a SRS of 11
three-year old children and obtains the following data. (See lecture notes for
data)
Begin by reading in the data:

```{r}
Height = c(27.75, 24.5, 25.5, 26, 25, 27.75, 26.5, 27, 26.75, 26.75, 27.5)
Circ = c(17.5, 17.1, 17.1,17.3, 16.9, 17.6, 17.3, 17.5, 17.3, 17.5, 17.5)
Dat = data.frame(Height,Circ)
attach(Dat)
Dat
```

To make a scatter plot of circumference against height type:
17.6
```{r}
plot(Dat$Height,Dat$Circ)
```

Studying the plot, there appears to be a linear relationship between the two
variables.
This relationship can be quantified by computing the covariance and correlation
between variables.

```{r}
cov(Dat) # Covariance matrix
```
From the output we see that the variance of Height and Circ is 1.198 and 0.048,
respectively. The covariance between the two variables is 0.219 indicating a
positive relationship.
```{r}
cor(Dat) # Correlation matrix
```

#Simple Linear Regression 

If there exists a strong linear relationship between two variables it is often of
interest to model the relationship using a regression line. The main function for
performing regression in R is lm(). It has many options that we will explore
throughout the semester.
To perform simple linear regression we can use the command:
lm(response ~ explanatory)
Here the terms response and explanatory in the function should be replaced by
the names of the response and explanatory variables, respectively, used in the
analysis.
Ex. Fit a regression line that describes the relationship between Height and
Circumference.
```{r}
results = lm(Circ ~ Height)
results
```

The results indicate that the least squares regression line takes the form: yhat =
12.493 + 0.183x. Hence the model states that a one inch increase in height
would lead to a 0.183 inch increase in head circumference.
To superimpose the regression line over the data first make a scatter plot of Circ
against Height, and therefore overlay the regression line using the command
abline(results). Here results contains all relevant information about the regression
line.
```{r}
plot(Height,Circ)
abline(results)
```
